---
title: "Statistical analysis ADC Study"
author: "Vitus Gosch"
date: '2022-11-17'
output: html_document
---

Df "Xrepro_annonymized" was created by reading in text file output out of shell scripts written to analyze image data. The Df is available in annonymized form, meaning all possible identifiers are removed.

Nomenclature: # importing data or packages   ## statistical analysis  ### visualization


*############# Load data and packages #############*

```{r}
# Load data
load(file = "Xrepro_annonymized.RData")
X <- Xrepro_annonymized

# Load packages
library(GGally)
library(ggplot2)

library(dplyr) 

library(viridis)
library(hrbrthemes)

library(tidyr)

library(broom)
library(forestmangr)

library(lmtest)
library(gvlma)

```


*################ Figure 3 ################*
*Using summary to assess volume and ADC distribution. Building boxplots and ggparcoord visuals*

```{r}

## Tests for normal distribution - Decide wether variables are normally distributed or not and for what reason you could make an exception
shapiro.test(X$VOL_1)
shapiro.test(X$VOL_manual_1)
shapiro.test(X$MEAN_auto_1)
shapiro.test(X$MEAN_manual_1)


## Day 1 comparison volume   
complete_VOL_evo <- X %>% select("VOL_1","VOL_manual_1")
complete_VOL_evo <- dplyr::rename(complete_VOL_evo, "Automated ROI" = VOL_1)
complete_VOL_evo <- dplyr::rename(complete_VOL_evo, "Manual ROI" = VOL_manual_1)
mean_VOL_auto <- mean(X$VOL_1)
mean_VOL_manual <- mean(X$VOL_manual_1)

summary(complete_VOL_evo$`Automated ROI`)
summary(complete_VOL_evo$`Manual ROI`)


## Day 1 comparison ADC  
complete_MEAN_evo <- X %>% select("MEAN_auto_1","MEAN_manual_1")
complete_MEAN_evo <- dplyr::rename(complete_MEAN_evo, "Automated ROI" = MEAN_auto_1)
complete_MEAN_evo <- dplyr::rename(complete_MEAN_evo, "Manual ROI" = MEAN_manual_1)
mean_MEAN_auto <- mean(X$MEAN_auto_1, na.rm = TRUE)
mean_MEAN_manual <- mean(X$MEAN_manual_1)

summary(complete_MEAN_evo$`Automated ROI`)
summary(complete_MEAN_evo$`Manual ROI`)



### ggparcord general appaerance
        
#png("descriptive_A.png", units="in", width=5, height=5, res=300)
 
ggparcoord (data=complete_MEAN_evo, columns = 1:2, scale="globalminmax", showPoints = TRUE, alphaLines = 0.1, boxplot = TRUE) +
    theme(legend.position="none", axis.title.x=element_blank(),
        axis.text.x=element_text(color = "black",size = 20),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")) +
        ylab(expression(paste("Mean ADC in 10"^"−6","mm"^"2","/s"))) + theme(text=element_text(size=15, family="Arial")) +
  labs(caption = "A") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))


#png("descriptive_B.png", units="in", width=5, height=5, res=300)

ggparcoord (data=complete_VOL_evo, columns = 1:2, scale="globalminmax", showPoints = TRUE, alphaLines = 0.1, boxplot = TRUE) +
    theme(legend.position="none", axis.title.x=element_blank(),
        axis.text.x=element_text(color = "black",size = 20),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")) +
        ylab("Volume (ml)") + theme(text=element_text(size=15, family="Arial")) +
  labs(caption = "B") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))


```


*################ Figure 4A ################ Bland Altman analysis*
*Generating new df “abs” and finding mean absolute differences, CIs. Performing BA analysis with plotting X instead of X1+X2/2 because X is a reference method. (https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.3086 Why Bland–Altman plots should use X, not (Y+X)/2 when X is a reference method).* 

```{r}
## Bland Altman plot against X - Try ABSOLUTE values
abs <- X %>% select(VOL_manual_1, VOL_1) %>% mutate(diff=VOL_1 - VOL_manual_1,          avg=(VOL_manual_1+VOL_1)/2)
abs$pdiff <- as.numeric((abs$diff / abs$VOL_manual_1))
    
## find average difference & 95% confidence interval limits
mean_diff <- mean(abs$diff)
mean_diff

lower <- mean_diff - 1.96*sd(abs$diff)
lower

upper <- mean_diff + 1.96*sd(abs$diff)
upper

### plot with x as reference method

#png("volumetrics_A_18.png", units="in", width=5, height=5, res=300)

ggplot(abs, aes(x = VOL_manual_1, y = diff)) +
  geom_point(size=2, alpha = 1) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = lower, color = "red", linetype="dashed") +
  geom_hline(yintercept = upper, color = "red", linetype="dashed") +
  ylab("Differences between methods (ml)") +
  xlab("Manual ROI (ml)") +
  coord_equal() +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 18),axis.text =     
            element_text(size = 20, colour = "black")) + 
  labs(caption = "A") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))

dev.off()

```


*################ Figure 4B ################ Volumetric analysis*
*Volumetric agreement. Log scale to also show differences in low volumes.*

```{r}


## Correlation and linear regression

cor.test(X$VOL_1, X$VOL_manual_1)
lm_vol <- lm(VOL_1~VOL_manual_1, data = X)
summary(lm_vol)


### Plot log comparison

#png("volumetrics_B.png", units="in", width=5, height=5, res=300)

ggplot(X, aes(VOL_manual_1, VOL_1)) + 
    geom_point( color = "black", size = 2) + labs(x = "Manual ROI (ml)", y = "Automated ROI (ml)") +       geom_smooth(method="lm", color = "blue") +
    scale_x_continuous(trans='log', breaks = c(0.1, 1.00,10.00,80.00)) + 
    scale_y_continuous(trans='log', breaks = c(0.1, 1.00,10.00,80.00)) + 
    coord_equal() +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title = element_text(size = 20), axis.text =     
            element_text(size = 20, colour = "black")) + 
  labs(caption = "B") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))


```


*################ Figure 4C ################ DICE coefficient*

```{r}

## DICE score

DSC <- select(X, "DSC_1")
summary(DSC)


### DICE visualization

#png("volumetrics_C.png", units="in", width=5, height=5, res=300)

DSC %>%
    gather(key="MesureType", value="Val") %>%
    ggplot(aes(x=MesureType, y=Val, fill=MesureType)) +
    geom_boxplot(fill = "white") +
    geom_jitter(color="black", size=2, alpha=1, width = 0.2) +
    theme(
        legend.position="none", axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")
    ) +
    ggtitle("") +
    ylab("Dice coefficient") + xlab("") + 
  labs(caption = "C") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))



```


*################ Figure 5 ################ ROC analysis*
*Only part were additional data was used distant from "X" df. Files to be read in are direct output of fslmaths - roc. TPR, FPR and ADC value were read in separately. Specificity was calculated. "output_roc_fast" files contain maximum output in the primary file (therefore not pooled results) but best possible threshold for each subject*  

```{r}

# load ROC files, priorly read in as txt files of 1001 line length

load(file = "TPR_all4.RData")
load(file = "TPR_all4_dil.RData")

load(file = "FPR_all4.RData")
load(file = "FPR_all4_dil.RData")

load(file = "ADC_all4.RData")
load(file = "ADC_all4_dil.RData")

load(file = "Spec_all4.RData")
load(file = "Spec_all4_dil.RData")

load(file = "Youden_all4.RData")
load(file = "Youden_all4_dil.RData")

load(file = "output_roc_fast4_repro_annonymized.RData")
load(file = "output_roc_fast4_dil_repro_annonymized.RData")


## Generate pooled data by adding up all rows horizontally 

FPR_all4_mean <- rowMeans(FPR_all4)
TPR_all4_mean <- rowMeans(TPR_all4)
ADC_all4_mean <- rowMeans(ADC_all4)
Youden_all4_mean <- rowMeans(Youden_all4)
FPR_all4_mean <- as.data.frame(FPR_all4_mean)
FPR_all4_mean$row <- seq.int(nrow(FPR_all4_mean))
TPR_all4_mean <- as.data.frame(TPR_all4_mean)
TPR_all4_mean$row <- seq.int(nrow(TPR_all4_mean))
ADC_all4_mean <- as.data.frame(ADC_all4_mean)
ADC_all4_mean$row <- seq.int(nrow(ADC_all4_mean))
Youden_all4_mean <- as.data.frame(Youden_all4_mean)
Youden_all4_mean$row <- seq.int(nrow(Youden_all4_mean))
all4_mean <- merge(FPR_all4_mean, TPR_all4_mean, by="row")
all4_mean <- merge(all4_mean, ADC_all4_mean, by="row")
all4_mean <- merge(all4_mean, Youden_all4_mean, by="row")

FPR_all4_dil_mean <- rowMeans(FPR_all4_dil)
TPR_all4_dil_mean <- rowMeans(TPR_all4_dil)
ADC_all4_dil_mean <- rowMeans(ADC_all4_dil)
Youden_all4_dil_mean <- rowMeans(Youden_all4_dil)
FPR_all4_dil_mean <- as.data.frame(FPR_all4_dil_mean)
FPR_all4_dil_mean$row <- seq.int(nrow(FPR_all4_dil_mean))
TPR_all4_dil_mean <- as.data.frame(TPR_all4_dil_mean)
TPR_all4_dil_mean$row <- seq.int(nrow(TPR_all4_dil_mean))
ADC_all4_dil_mean <- as.data.frame(ADC_all4_dil_mean)
ADC_all4_dil_mean$row <- seq.int(nrow(ADC_all4_dil_mean))
Youden_all4_dil_mean <- as.data.frame(Youden_all4_dil_mean)
Youden_all4_dil_mean$row <- seq.int(nrow(Youden_all4_dil_mean))
all4_dil_mean <- merge(FPR_all4_dil_mean, TPR_all4_dil_mean, by="row")
all4_dil_mean <- merge(all4_dil_mean, ADC_all4_dil_mean, by="row")
all4_dil_mean <- merge(all4_dil_mean, Youden_all4_dil_mean, by="row")

## Pooled ROC analysis result, max. YI, sens + spec + AUC

all4_mean[which.max(all4_mean$Youden_all4_mean),]
all4_dil_mean[which.max(all4_dil_mean$Youden_all4_dil_mean),]
library(pracma)
trapz(FPR_all4_mean[,1], TPR_all4_mean[,1])
trapz(FPR_all4_dil_mean[,1], TPR_all4_dil_mean[,1])
    "AUC will be inverted due to reversed order"


########################### Viszualize pooled ROC data #####################


### Plot all mean ROC curves  

plot(rowMeans(FPR_all4), rowMeans(TPR_all4), type = "l", lwd = "3", lty = 1, col = 1, add=TRUE, xlim = c(0,1), ylim = c(0,1), xlab = "FPR", ylab = "TPR", cex.lab = 1.6, cex.axis = 1.6)
lines(rowMeans(FPR_all4_dil), rowMeans(TPR_all4_dil), type = "l", lwd = "3", lty = 1, col = 3)
abline(a=0, b=1, lty = 3)

plot(rowMeans(ADC_all4), rowMeans(Youden_all4), type = "l", lwd = "3", lty = 1, col = 1, add=TRUE, xlim = c(0,2500), ylim = c(0,0.5), xlab = "Pooled ADC Thresholds (x10^−6 mm2/s)", ylab = "Pooled Youden Index", cex.lab = 1.3, cex.axis = 1.3)
lines(rowMeans(ADC_all4_dil), rowMeans(Youden_all4_dil), type = "l", lwd = "3", lty = 1, col = 3)
legend(1500, 0.5, legend=c("Whole brain", "Dilated DWI lesion"),
       col=c("black", "green"), lty=1, cex=1, bty = "n")

### Individual and mean Curves combined 

#png("ROC_A.png", units="in", res=300, width = 5, height = 5)

matplot(FPR_all4, TPR_all4,  xaxs = "i", yaxs = "i", type = "l", lwd = "0.5", lty = 1, col = rgb(0,0,0, alpha = 0.3), xlab = "FPR", ylab = "TPR", main = "", cex = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.4)
matplot(FPR_all4_dil, TPR_all4_dil,  xaxs = "i", yaxs = "i", type = "l", lwd = "0.5", lty = 1, col = rgb(0,0,1, alpha = 0.2), add=TRUE)
abline(a=0, b=1, lty=3)
matplot(rowMeans(FPR_all4), rowMeans(TPR_all4), type = "l", lwd = "3", lty = 1, col = rgb(0,0,0), add=TRUE)
matplot(rowMeans(FPR_all4_dil), rowMeans(TPR_all4_dil), type = "l", lwd = "3", lty = 1,  col = rgb(0,0,1), add=TRUE) 
mtext("A", side = 1, line = 4, cex = 1.5, adj = -0.2) 


#png("ROC_B.png", units="in", res=300, width = 5, height = 5)

lim <- c(0,2000)
matplot(ADC_all4, Youden_all4,xaxs = "i", yaxs = "i" , type = "l", lwd = "0.5", lty = 1, col = rgb(0,0,0, alpha = 0.15), xlab = expression(paste("Mean ADC in 10"^"−6","mm"^"2","/s")), ylab = "Youden Index", main = "", cex = 1.3, xlim = lim, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)
matplot(ADC_all4_dil, Youden_all4_dil,xaxs = "i", yaxs = "i" , type = "l", lwd = "0.5", lty = 1, col = rgb(0,0,1, alpha = 0.12), add = TRUE)
matplot(rowMeans(ADC_all4), rowMeans(Youden_all4), type = "l", lwd = "3", lty = 1,col = 1, add=TRUE)
matplot(rowMeans(ADC_all4_dil), rowMeans(Youden_all4_dil), type = "l", lwd = "3", lty = 1, col = rgb(0,0,1), add=TRUE)
legend(850, 0.85, legend=c("Whole brain", "Dilated DWI lesion"),
       col=c("black", "blue"), lty=1, cex=1, lwd = 2, bty = "n") 
mtext("B", side = 1, line = 4, cex = 1.5, adj = -0.2) 


```


*################ Figure 6 ################ Linear Model*
*Using lm as a function, assessing assumptions via gvlma and visually*

```{r}

## Plain Approach standardized - Exclude wake up strokes
Xs <- X[,c("DSC_1", "TOO", "age", "MEAN_manual_1", "VOL_manual_1", "MEAN_fast", "youden_fast4")]
Xs <- Xs[Xs$TOO != 720, ]
Xs <- Xs %>% mutate_at(c("TOO", "age", "MEAN_manual_1", "VOL_manual_1", "MEAN_fast", "youden_fast4", "DSC_1"), scale)

lmyou <- lm(formula = Xs$youden_fast4 ~ Xs$TOO + Xs$age + Xs$VOL_manual_1 + Xs$MEAN_manual_1 + Xs$MEAN_fast, data = Xs)
names(lmyou$coefficients) <- c("(Intercept)", "TOO","Age","Volume DWI Lesion","Mean ADC DWI Lesion", "Mean ADC Parenchyma")

lmdsc <- lm(formula = Xs$DSC_1 ~ Xs$TOO + Xs$age + Xs$VOL_manual_1 + Xs$MEAN_manual_1 +  Xs$MEAN_fast, data = Xs)
names(lmdsc$coefficients) <- c("(Intercept)", "TOO","Age","Volume DWI Lesion","Mean ADC DWI Lesion", "Mean ADC Parenchyma")

summary(lmyou)
summary(lmdsc)


## Testing assumptions

gvlma::gvlma.lm(lmobj = lmdsc) 
gvlma::gvlma.lm(lmobj = lmyou) 

par(mfrow = c(2, 2))
plot(lmdsc)
par(mfrow = c(2, 2))
plot(lmyou)


### Plot both models

                             "DICE"
coef <- coef(lmdsc)
ConfidenceInterval <- confint(lmdsc)
coef_confint <- cbind(coef, ConfidenceInterval) %>% as.data.frame()
coef_confint <- coef_confint %>% mutate(variable=rownames(coef_confint))
library(plyr) 
coef_confint <- rename(coef_confint,c("coef" = "Beta",
                                      `2.5 %` = "lower_bound", 
                                      `97.5 %` = "upper_bound"))

# We don't need to plot the intercept. We can remove it from our data. Reorder variables:
col_order <- c("variable", "Beta", "lower_bound", "upper_bound")
coef_confint <- coef_confint[, col_order] #reorder variables in the data frame

plot_lm_dsc <- coef_confint[-1,] %>%  #remove row number 1 (The intercept) 
  ggplot(aes(x=reorder(variable, Beta), y=Beta)) +
  geom_point(shape = 16,
             size  = 3, width = 0.2,
             position = "dodge", color="black") + 
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = .2,
                size  = 0.8,
                position = "dodge", color="black") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Predictor Variables") + ylab("Beta (Response = Dice)") + 
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) + 
  coord_flip(ylim = c(-1, 1)) + 
  geom_hline(yintercept = 0, color = "black", size = 0.5, linetype = "dashed") +
  theme(axis.title = element_text(size = 35)) + 
  theme(axis.text = element_text(size = 35, colour = "black")) +
  theme_bw(base_size = 35) + 
  theme(panel.background = (element_blank())) + 
  labs(caption = "A") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 35))


                                   "Youden Index"
coef <- coef(lmyou)
ConfidenceInterval <- confint(lmyou)
coef_confint <- cbind(coef, ConfidenceInterval) %>% as.data.frame()
coef_confint <- coef_confint %>% mutate(variable=rownames(coef_confint))
library(plyr) 
coef_confint <- rename(coef_confint,c("coef" = "Beta",
                                      `2.5 %` = "lower_bound", 
                                      `97.5 %` = "upper_bound"))

# We don't need to plot the intercept. We can remove it from our data. Reoiorder variables:
col_order <- c("variable", "Beta", "lower_bound", "upper_bound")
coef_confint <- coef_confint[, col_order] #reorder variables in the data frame

plot_lm_you <- coef_confint[-1,] %>%  #remove row number 1 (The intercept) 
  ggplot(aes(x=reorder(variable, Beta), y=Beta)) +
  geom_point(shape = 16,
             size  = 3, width = 0.2,
             position = "dodge", color="black") + 
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = .2,
                size  = 0.8,
                position = "dodge", color="black") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Predictor Variables") + ylab("Beta (Response = YI)") + 
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) + 
  coord_flip(ylim = c(-1, 1)) + 
  geom_hline(yintercept = 0, color = "black", size = 0.5, linetype = "dashed") +
  theme(axis.title = element_text(size = 35)) + 
  theme(axis.text = element_text(size = 35, colour = "black")) +
  theme_bw(base_size = 35) + 
  theme(panel.background = (element_blank())) + 
  labs(caption = "B") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 35))
  

png("lm_A.png", units="in", res=300, width = 10, height = 10)
plot_lm_dsc
dev.off()

png("lm_B.png", units="in",  res=300, width = 10, height = 10)
plot_lm_you
dev.off()

```


