---
title: "Statistical analysis ADC Study - Supplementary"
author: "Vitus Gosch"
date: '2022-11-17'
output: html_document
---

Df "Xrepro_annonymized" was created by reading in text file output out of shell scripts written to analyze image data. The Df is available in annonymized form, meaning all possible identifiers are removed.

Nomenclature: # importing data or packages   ## statistical analysis  ### visualization


*############# Load data and packages #############*

```{r}
# Load data
load(file = "Xrepro_annonymized.RData")
X <- Xrepro_annonymized

# Load packages
library(GGally)
library(ggplot2)

library(dplyr) 

library(viridis)
library(hrbrthemes)

library(tidyr)

library(broom)
library(forestmangr)

library(lmtest)
library(gvlma)

```


*################ Table S1 and figure S1 ################ *

```{r}

### DICE score against volume

ggplot(X, aes(VOL_manual_1, DSC_1)) + 
    geom_point( color = "black", size = 3) + labs(x = "Manual ROI (cm³)", y = "DICE") +       geom_smooth(method="lm", color = "black") +
    scale_x_continuous(breaks = c(0,5,15,30,75)) + 
    scale_y_continuous(limits=c(0,1)) + 
    theme(aspect.ratio=1) +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title = element_text(size = 25), axis.text =     
              element_text(size = 25, colour = "black")) 


## Pearson product correlation
       
cor.test(X$youden_fast4, X$age)
cor.test(X$youden_fast4, X$VOL_manual_1)

cor.test(Xs$youden_fast4, Xs$age)
cor.test(Xs$youden_fast4, Xs$VOL_manual_1)


```


*################ Figures S2 and S3 ################ *
*Univariate Linear Models*


```{r}

## Plain Approach standardized - Exclude wake up strokes
Xs <- X[,c("DSC_1", "TOO", "age", "MEAN_manual_1", "VOL_manual_1", "MEAN_fast", "youden_fast4")]
Xs <- Xs[Xs$TOO != 720, ]
Xs <- Xs %>% mutate_at(c("TOO", "age", "MEAN_manual_1", "VOL_manual_1", "MEAN_fast", "youden_fast4", "DSC_1"), scale)

lmyou <- lm(formula = Xs$youden_fast4 ~ Xs$TOO + Xs$age + Xs$VOL_manual_1 + Xs$MEAN_manual_1 + Xs$MEAN_fast, data = Xs)
names(lmyou$coefficients) <- c("(Intercept)", "TOO","Age","Volume DWI Lesion","Mean ADC DWI Lesion", "Mean ADC Parenchyma")

lmdsc <- lm(formula = Xs$DSC_1 ~ Xs$TOO + Xs$age + Xs$VOL_manual_1 + Xs$MEAN_manual_1 +  Xs$MEAN_fast, data = Xs)
names(lmdsc$coefficients) <- c("(Intercept)", "TOO","Age","Volume DWI Lesion","Mean ADC DWI Lesion", "Mean ADC Parenchyma")

summary(lmyou)
summary(lmdsc)


## Testing assumptions

gvlma::gvlma.lm(lmobj = lmdsc) 
gvlma::gvlma.lm(lmobj = lmyou) 

par(mfrow = c(2, 2))
plot(lmdsc)
par(mfrow = c(2, 2))
plot(lmyou)


## → univariate models (e.g. in supplementary mats)  interpretation (algorithm “better” vs “good”)
## + mean ADC DWI lesion effect adjusted for other predictors for the multiple regression model 


DSC_TOOlm <- lm(Xs$DSC_1~Xs$TOO, data = Xs)

# Show attributes of linModel
attributes(DSC_TOOlm) 
DSC_TOOlm$na.action
DSC_TOOlm$coefficients
summary(DSC_TOOlm)

# plot
png("DSC_1~Xs$TOO.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$DSC_1~Xs$TOO, data = Xs,
       xlab = "TOO (in minutes)",
       ylab = "Dice coefficient",
       main = "Univariate Regression",
       sub = expression(paste("Adjusted R"^"2",": -0.01111, p-value: 0.9819"))) +
abline(DSC_TOOlm, col = "red", lwd = 3) 
dev.off()


DSC_agelm <- lm(Xs$DSC_1~Xs$age, data = Xs)

# Show attributes of linModel
attributes(DSC_agelm) 
DSC_agelm$na.action
DSC_agelm$coefficients

# plot
png("DSC_1~Xs$age.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$DSC_1~Xs$age, data = Xs,
       xlab = "Age",
       ylab = "Dice coefficient",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.08989, p-value: 0.00214"))) +
abline(DSC_agelm, col = "red", lwd = 3) 
dev.off()



DSC_VOL_manual_1lm <- lm(Xs$DSC_1~Xs$VOL_manual_1, data = Xs)

# Show attributes of linModel
attributes(DSC_VOL_manual_1lm) 
DSC_VOL_manual_1lm$na.action
DSC_VOL_manual_1lm$coefficients

# plot
png("DSC_1~Xs$VOL_manual_1.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$DSC_1~Xs$VOL_manual_1, data = Xs,
       xlab = "Manual ROI (ml)",
       ylab = "Dice coefficient",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.1021, p-value: 0.00111"))) +
abline(DSC_agelm, col = "red", lwd = 3) 
dev.off()



DSC_MEAN_manual_1lm <- lm(Xs$DSC_1~Xs$MEAN_manual_1, data = Xs)

# Show attributes of linModel
attributes(DSC_MEAN_manual_1lm) 
DSC_MEAN_manual_1lm$na.action
DSC_MEAN_manual_1lm$coefficients

# plot
png("DSC_1~Xs$MEAN_manual_1.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$DSC_1~Xs$MEAN_manual_1, data = Xs,
       xlab = (expression(paste("Mean ADC manual ROI (10"^"−6","mm"^"2","/s)"))),
       ylab = "Dice coefficient",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.4501, p-value: <0.001"))) +
abline(DSC_MEAN_manual_1lm, col = "red", lwd = 3) 
dev.off()


DSC_MEAN_fastlm <- lm(Xs$DSC_1~Xs$MEAN_fast, data = Xs)

# Show attributes of linModel
attributes(DSC_MEAN_fastlm) 
DSC_MEAN_fastlm$na.action
DSC_MEAN_fastlm$coefficients

# plot
png("DSC_1~Xs$MEAN_fast.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$DSC_1~Xs$MEAN_fast, data = Xs,
       xlab = (expression(paste("Mean ADC Parenchyma (10"^"−6","mm"^"2","/s)"))),
       ylab = "Dice coefficient",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.01551, p-value: 0.1223"))) +
abline(DSC_MEAN_fastlm, col = "red", lwd = 3) 
dev.off()

plot(Xs$DSC_1~Xs$MEAN_fast, data = Xs,
       xlab = "MEAN_fast",
       ylab = "DSC",
       main = "Scatter Plot") +
abline(DSC_MEAN_fastlm, col = "red", lwd = 3)

tidy(DSC_TOOlm)
tidy(DSC_agelm)
tidy(DSC_VOL_manual_1lm)
tidy(DSC_MEAN_manual_1lm)
tidy(DSC_MEAN_fastlm)


## youden
## → univariate models (e.g. in supplementary mats)  interpretation (algorithm “better” vs “good”)
## + mean ADC DWI lesion effect adjusted for other predictors for the multiple regression model 


youden_TOOlm <- lm(Xs$youden_fast4~Xs$TOO, data = Xs)

# Show attributes of linModel
attributes(youden_TOOlm) 
youden_TOOlm$na.action
youden_TOOlm$coefficients

# plot
png("youden_fast4$TOO.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$youden_fast4~Xs$TOO, data = Xs,
       xlab = "TOO (in minutes)",
       ylab = "Youden Index",
       main = "Univariate Regression",
       sub = expression(paste("Adjusted R"^"2",": -0.008069, p-value: 0.6035"))) +
abline(youden_TOOlm, col = "red", lwd = 3) 
dev.off()


youden_agelm <- lm(Xs$youden_fast4~Xs$age, data = Xs)

# Show attributes of linModel
attributes(youden_agelm) 
youden_agelm$na.action
youden_agelm$coefficients

# plot
png("youden_fast4$age.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$youden_fast4~Xs$age, data = Xs,
       xlab = "Age",
       ylab = "Youden Index",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.06914, p-value: 0.0065"))) +
abline(youden_agelm, col = "red", lwd = 3) 
dev.off()



youden_VOL_manual_1lm <- lm(Xs$youden_fast4~Xs$VOL_manual_1, data = Xs)

# Show attributes of linModel
attributes(youden_VOL_manual_1lm) 
youden_VOL_manual_1lm$na.action
youden_VOL_manual_1lm$coefficients

# plot
png("youden_fast4$VOL_manual_1.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$youden_fast4~Xs$VOL_manual_1, data = Xs,
       xlab = "Manual ROI (ml)",
       ylab = "Youden Index",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": -0.01011, p-value: 0.7655"))) +
abline(youden_VOL_manual_1lm, col = "red", lwd = 3) 
dev.off()



youden_MEAN_manual_1lm <- lm(Xs$youden_fast4~Xs$MEAN_manual_1, data = Xs)

# Show attributes of linModel
attributes(youden_MEAN_manual_1lm) 
youden_MEAN_manual_1lm$na.action
youden_MEAN_manual_1lm$coefficients

# plot
png("youden_fast4$MEAN_manual_1.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$youden_fast4~Xs$MEAN_manual_1, data = Xs,
       xlab = (expression(paste("Mean ADC manual ROI (10"^"−6","mm"^"2","/s)"))),
       ylab = "Youden Index",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.8135, p-value: <0.001"))) +
abline(youden_MEAN_manual_1lm, col = "red", lwd = 3) 
dev.off()


youden_MEAN_fastlm <- lm(Xs$youden_fast4~Xs$MEAN_fast, data = Xs)

# Show attributes of linModel
attributes(youden_MEAN_fastlm) 
youden_MEAN_fastlm$na.action
youden_MEAN_fastlm$coefficients

# plot
png("youden_fast4$MEAN_fast.png", units="in",  res=500, width = 3.5, height = 3.5)
par(pty="s")
plot(Xs$youden_fast4~Xs$MEAN_fast, data = Xs,
       xlab = (expression(paste("Mean ADC Parenchyma (10"^"−6","mm"^"2","/s)"))),
       ylab = "Youden Index",
       main = "",
       sub = expression(paste("Adjusted R"^"2",": 0.005892, p-value: 0.2179"))) +
abline(youden_MEAN_fastlm, col = "red", lwd = 3) 
dev.off()

plot(Xs$youden_fast4~Xs$MEAN_fast, data = Xs,
       xlab = "MEAN_fast",
       ylab = "youden",
       main = "Scatter Plot") +
abline(youden_MEAN_fastlm, col = "red", lwd = 3)

tidy(youden_TOOlm)
tidy(youden_agelm)
tidy(youden_VOL_manual_1lm)
tidy(youden_MEAN_manual_1lm)
tidy(youden_MEAN_fastlm)

```



*################ Figure S4 ################ *
*FLAIR d6, infarct growth*

```{r}

# Get data 

X1 <- merge(X, FLAIR_d6, by = "date")

abs1 <- X1 %>% select(VOL_manual_1, VOL_1, FLAIR_d6) 
cc <- complete.cases(abs1)
abs1 <- abs1[cc,]

## Calculate differences d1 and d6 Flair volumes FLAIR as reference

        'analog bland altman'
abs1$diff_m <- as.numeric((abs1$FLAIR_d6 - abs1$VOL_manual_1 ))        
abs1$diff_a <- as.numeric((abs1$FLAIR_d6 - abs1$VOL_1  ))
    
## find average difference & 95% confidence interval limits
mean_diff_m <- mean(abs1$diff_m)
mean_diff_m
mean_diff_a <- mean(abs1$diff_a)
mean_diff_a

lower_m <- mean_diff_m - 1.96*sd(abs1$diff_m)
lower_m
upper_m <- mean_diff_m + 1.96*sd(abs1$diff_m)
upper_m

lower_a <- mean_diff_a - 1.96*sd(abs1$diff_a)
lower_a
upper_a <- mean_diff_a + 1.96*sd(abs1$diff_a)
upper_a

abs_m <- abs1$diff_m
ROI <- c(replicate(78,"Manual ROI"))
abs_m <- data.frame(abs_m, ROI)
abs_m$d6 <- abs1$FLAIR_d6
abs_a <- abs1$diff_a
ROI <- c(replicate(78,"Automated ROI"))
abs_a <- data.frame(abs_a, ROI)
abs_a$d6 <- abs1$FLAIR_d6
names(abs_a)[1] <- 'adiff'
names(abs_m)[1] <- 'adiff'
adiff <- rbind(abs_m, abs_a)
adiff$ROI <- as.character(adiff$ROI)

### plot with FLAIR as reference method 





png("FLAIR_ref.png", units="in",  res=300, width = 8, height = 8)

ggplot(adiff, aes(x = d6, y = adiff, color = ROI)) +
  geom_point(size=3, alpha = 0.8) +
  scale_color_manual(values = c("Manual ROI" = "black", "Automated ROI" = "blue")) +
  geom_hline(yintercept = mean_diff_m, color = "black") +
  geom_hline(yintercept = lower_m, color = "black", linetype="dashed") +
  geom_hline(yintercept = upper_m, color = "black", linetype="dashed") +
  geom_hline(yintercept = mean_diff_a, color = "blue") +
  geom_hline(yintercept = lower_a, color = "blue", linetype="dashed") +
  geom_hline(yintercept = upper_a, color = "blue", linetype="dashed") +
  ylab("Differences to day 1 (ml)") +
  xlab("Day 3-5 FLAIR (ml)") +
  scale_x_continuous() + 
  scale_y_continuous() +
  coord_equal() +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title = element_text(size = 30), axis.text =     
            element_text(size = 30, colour = "black"), legend.text = element_text(size = 30), legend.title = element_blank(), legend.position = "top") + 
  labs(caption = "A") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 30))


dev.off()


ggplot(abs1, aes(x = FLAIR_d6, y = diff_m)) +
  geom_point(size=2, alpha = 1) +
  geom_hline(yintercept = mean_diff_m, color = "blue") +
  geom_hline(yintercept = lower_m, color = "red", linetype="dashed") +
  geom_hline(yintercept = upper_m, color = "red", linetype="dashed") +
  ylab("Infarct Growth Day 6 (cm³)") +
  xlab("Manual ROI (cm³)") +
  scale_x_continuous(breaks = c(0,25,50,75,100)) + 
  scale_y_continuous(breaks = c(0,50,100)) +
  coord_equal() +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title = element_text(size = 20), axis.text =     
            element_text(size = 20, colour = "black"))

ggplot(abs1, aes(x = VOL_1, y = diff_a)) +
  geom_point(size=2, alpha = 1) +
  geom_hline(yintercept = mean_diff_a, color = "blue") +
  geom_hline(yintercept = lower_a, color = "red", linetype="dashed") +
  geom_hline(yintercept = upper_a, color = "red", linetype="dashed") +
  ylab("Infarct Growth (cm³)") +
  xlab("Automated ROI (cm³)") +
  scale_x_continuous(breaks = c(0,25,50,75,100)) + 
  scale_y_continuous(breaks = c(0,50,100)) +
  coord_equal() +
    theme_bw() +
    theme(panel.border = element_blank(), axis.title = element_text(size = 20), axis.text =     
            element_text(size = 20, colour = "black"))


### plot as boxplot

FLAIR <- select(X1,"VOL_1", "VOL_manual_1","FLAIR_d6")
FLAIR <- dplyr::rename(FLAIR, "Automated" = VOL_1)
FLAIR <- dplyr::rename(FLAIR, "Manual" = VOL_manual_1)
FLAIR <- dplyr::rename(FLAIR, "FLAIR day 6" = FLAIR_d6)


ggparcoord (data=FLAIR, columns = 1:3, scale="globalminmax", showPoints = TRUE, alphaLines = 0.1, boxplot = TRUE) +
    theme(legend.position="none", axis.title.x=element_blank(),
        axis.text.x=element_text(color = "black",size = 20),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")) +
        ylab("Volume in ml") + theme(text=element_text(size=15, family="Arial")) +
  labs(caption = "A") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))

### infarct growth

pabs1 <- abs1 %>% select(pdiff_a, pdiff_m) 

pabs1 %>%
    gather(key="MesureType", value="Val") %>%
    ggplot(aes(x=MesureType, y=Val, fill=MesureType)) +
    geom_boxplot(fill = "white") +
    geom_jitter(color="black", size=2, alpha=1, width = 0.2) +
    theme(
        legend.position="none", axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")
    ) +
    ggtitle("") +
    ylab("Dice coefficient") + xlab("") + 
  labs(caption = "C") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 20))



# Highlight -> loosing consecutive info

FLAIR <- select(X1,"VOL_1", "VOL_manual_1","FLAIR_d6")
FLAIR <- dplyr::rename(FLAIR, "Automated" = VOL_1)
FLAIR <- dplyr::rename(FLAIR, "Manual" = VOL_manual_1)
FLAIR <- dplyr::rename(FLAIR, "FLAIR day 6" = FLAIR_d6)
cc <- complete.cases(FLAIR)
FLAIR <- FLAIR[cc,]


Volume <- FLAIR$Manual
ID <- 1:78
Day <- c(replicate(78,"1"))
ROI <- c(replicate(78,"Manual"))
man <- data.frame(Volume, Day, ROI, ID)
Volume <- FLAIR$Automated
ROI <- c(replicate(78,"Automated"))
auto <- data.frame(Volume, Day, ROI, ID)
Volume <- FLAIR$FLAIR
Day <- c(replicate(78,"3-5"))
ROI <- c(replicate(78,"Flair"))
flair <- data.frame(Volume, Day, ROI, ID)
day <- rbind(man,auto,flair)

png("FLAIR_box.png", units="in",  res=300, width = 5, height = 5)

ggplot(day, aes(x=reorder(ROI,Volume), y=Volume, fill=Day)) + 
    geom_boxplot(outlier.alpha = 0) + geom_jitter(width = 0.13, size=1.5) + theme(
        legend.position="top", legend.text = element_text(color = "black",size = 20), legend.title = element_text(color = "black",size = 20),
        axis.title.x=element_blank(),
        axis.text.x=element_text(color = "black",size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(color = "black",size = 20),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "light grey"),
        panel.grid.major = element_line(colour = "light grey")
    ) + ylab("Volume (ml)") + xlab(c("Automated", "Manual", "Flair")) + 
    labs(caption = "B") +
    theme(plot.caption.position = "plot",
          plot.caption = element_text(hjust = 0, size = 20)) + scale_fill_manual(values=c("white", "gray"))

dev.off()

```


